---
title: "python4"
author: "Levi Orero"
date: "5/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import library
## R
```{r, message=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
```

## Python
```{python}
import pandas as pd
import numpy as np
import seaborn as sns
import warnings
warnings.filterwarnings("ignore")
import matplotlib.pyplot as plt
sns.set(style = "white", color_codes = True)

```

# Creating a DataFrame
## R
```{r}
dates <- seq(as.Date("2013/01/01"), by = "day", length.out = 6)

df <- data.frame(date = dates, 
                 A = runif(6),
                 B = runif(6),
                 C = runif(6),
                 D = runif(6))

df
```

## Python
```{python}
dates = pd.date_range('20130101',periods = 6)
df = pd.DataFrame(np.random.randn(6,4), index = dates, columns = list('ABCD'))
df
```

## R
```{r}
dates <- as.Date("2013/01/01")
df2 <- data.frame(date = dates,
                  A = 1,
                  B = runif(4),
                  C = runif(4),
                  D = runif(4),
                  E = c("test", "train","test","train"),
                  F = "foo")
df2
```

## Python
```{python}
df2 = pd.DataFrame({
'A':1.,
'B':pd.Timestamp('20130102'),
'C':pd.Series(1,index=list(range(4)),dtype='float32'),
'D':np.array([3]*4,dtype='int32'),
'E':pd.Categorical(["test","train","test","train"]),
'F':'foo'})
df2
```

## R
```{r}
sapply(df, class)
```

## Python
```{python}
df2.dtypes
```

# Viewing Data
## R
```{r}
head(df)
tail(df,3)
dim(df)
nrow(df)
ncol(df)
```
## Python
```{python}
df.head()
df.tail(3)
df.shape

```
```{python}
df.columns
```

## R
```{r}
summary(df)
```

## Python
```{python}
#Describe shows a quick summary of your data
df.describe()
```

## R
```{r}
t(df)
```

## Python
```{python}
df.T
```

## R
```{r}
df[order(df$B),] #sort values
```

## Python
```{python}
df.sort_values(by='B')
```

# Data Manipulation
## R
```{r}
df[,c('A')]
```

## Python
```{python}
df['A']
```

## R
```{r}
df[c(1:3),]
```
## Python
```{python}
df[0:3]
```

## R
```{r}
df[df$date >= '2013-01-02' & df$date <= '2013-01-04',]
```
## Python
```{python}
df['20130102':'20130104']
```

# Selection by Label
## R
```{r}
df[,c('A','B')]
```
## Python
```{python}
df.loc[:,['A','B']]
```

## R
```{r}
df[df$date >= '2013-01-02' & df$date <= '2013-01-04', c('A','B')]
```

## Python
```{python}
df.loc['20130102':'20130104',['A','B']]
```

## R
```{r}
df[(df$date == '2013-01-02'), c('A','B')]
```

## Python
```{python}
df.loc['20130102',['A','B']]
```

# Selection by Position
## R
```{r}
df[3,]
```
## Python
```{python}
df.iloc[3]
```

## R
```{r}
df[3:5,1:3]
```

## Python
```{python}
df.iloc[2:5,0:2]
```

## R
```{r}
df[c(2,3,5),c(1,3)]
```

## Python
```{python}
df.iloc[[1,2,4],[0,2]]
```
## R
```{r}
df[c(2:4),]
```

## Python
```{python}
df.iloc[1:3,:]
```

## R
```{r}
df[,c(2:4)]
```
## Python
```{python}
df.iloc[:,1:3]
```

## R
```{r}
df[2,2]
```
## Python
```{python}
df.iloc[1,1]
```

## R
```{r}
df[2,2]
```
## Python
```{python}
df.iat[1,1]
```

## R
```{r}
df[df$A>0,]
```
## Python
```{python}
df[df.B > 0.8]
```

## R
```{r}
df2 <- df
df2$E <- c('one','one','two','three','four','three')
df2
```
## Python
```{python}
df2 = df.copy()
df2['E'] = ['one', 'one','two', 'three','four','three']
df2
```

# Dealing with Missing Data
## R
```{r}
dates <- seq(
  as.Date("2013/01/01"),
  by = "day",
  length.out = 6)

df1 <- data.frame(date = dates[1:4])
df1 <- cbind(df1,df[1:4,c('A','B','C','D')])
df1$E <- 1
df1[3:4,'E'] <- NA
df1

```
## Python
```{python}
df1 = df.reindex(index = dates[0:4],
columns = list(df.columns) + ['E'])

df1.loc[dates[0]:dates[1],'E'] = 1
df1
```

## R
```{r}
#Drop missing data in R
na.omit(df1)
```

## Python
```{python}
df1.dropna(how = 'any')
```

## R - Fill in missing data
```{r}
df2 <- df1
df1[is.na(df1)] <- 5
df1
```
## Python
```{python}
df1.fillna(value=3)
```

## R
```{r}
is.na(df2)
```
## Python
```{python}
pd.isnull(df1)
```

# Statistical Operations
## R
```{r}
sapply(df[,-1],mean, na.rm = T)
```
## Python
```{python}
df.mean()
```
## R
```{r}
data.frame(ID = df[,1], Means = rowMeans(df[,-1]))
```
## Python
```{python}
df.mean(1)
```

# String Methods
## R
```{r}
s <- c('A','B','C','Aaba','Baca',NA,'CABA','dog','cat')
tolower(s)
```

## Python
```{python}
s = pd.Series(['A','B','C','Aaba','Baca',np.nan,'CABA','dog','cat'])
s.str.lower()
```
# Data frames Merge operation
## R
```{r}
df <- data.frame(
  A = runif(10),
  B = runif(10),
  C = runif(10),
  D = runif(10))
df

```
## Python
```{python}
df = pd.DataFrame(np.random.randn(10,4))
df
```
## R
```{r}
a <- df[1:3,]
b <- df[3:7,]
c <- df[7:10,]

rbind(a,b,c)
```
## Python
```{python}
pieces = [df[:3],df[3:7],df[7:]]
pd.concat(pieces)
df[:3]
df[3:7]
pieces
```

# Join data frames
## R
```{r}
left <- data.frame(key = 'foo',
                   lval = c(1,2))
right <- data.frame(key = 'foo',
                    lval = c(4,5))
left
right
```
## Python
```{python}
left = pd.DataFrame({'key': ['foo','foo'],'lval': [1,2]})
right = pd.DataFrame({'key':['foo','foo'],'rval':[4,5]})
```

## R
```{r}
merge(x = left, y = right, by = "key", all = TRUE)
```

## Python
```{python}
pd.merge(left, right, on = 'key')
```

# Import dataset
```{r}
data <- read_csv('boston_test_data.csv')
```
## Python
```{python}
data_p = pd.read_csv('boston_test_data.csv')
```

## R
```{r}
str(data)
```
## Python
```{python}
data_p.info()
```
## R
```{r}
colnames(data)
```

## Python
```{python}
data_p.columns
```

## R
```{r}
head(data)
```
## Python
```{python}
data_p.head(6)
```

## R
```{r}
series <- data[,c('indus')]
print(class(series))
```
## Python
```{python}
series = data_p['indus']
print(type(series))
```

## R
```{r}
x = data[data$zn > 70,]
x
```

## Python
```{python}
x = data_p['zn']>70
data_p[x]
```
## R
```{r}
#Filtering with logic in R
x = data[data$zn > 70 & data$age <30,]
x
```
## Python
```{python}
data_p[np.logical_and(data_p['zn']>70,data_p['age']<30)]
data_p[(data_p['zn']>70) & (data_p['age']<30)]
```

# Exploratory Data Analysis
## R
```{r}
data %>% 
  group_by(rad) %>% 
  summarise(count = n())
```

## Python
```{python}
print(data_p['rad'].value_counts(dropna = False))
data_p.describe()
```

# CONCATENATING DATA
## R
```{r}
data1 <- head(data)
data2 <- tail(data)
conc_data_row <- rbind(data1,data2)
conc_data_row
```
## Python
```{python}
data1 = data_p.head()
data2 = data_p.tail()
conc_data_row = pd.concat([data1,data2],axis = 0, ignore_index = True)
conc_data_row
```
##R
```{r}
data1 <- head(data$crim)
data2 <- head(data$indus)
conc_data_col <- cbind(data1,data2)
conc_data_col
```
##Python
```{python}
data1 = data['crim'].head()
data2 = data['indus'].head()
conc_data_col = pd.concat([data1,data2],axis = 1)
conc_data_col
```
##R
```{r}
data[2,c("tax")]
```
##Python
```{python}
data["tax"][0]
data.loc[1,['tax']]
```
##R
```{r}
#Select specific columns
data[,c("tax","indus")]
```
##Python
```{python}
data[["tax","indus"]]
```
##R
```{r}
data[1:10,c("tax","indus")]
```
##Python
```{python}
data.loc[1:10,"tax":"black"]
```
##R
```{r}
data[1:10,9:11]
```
##Python
```{python}
data.loc[1:10,"rad":"black"]
```

# FILTERING DATA FRAMES
##R
```{r}
data %>%
  filter(rad > 2)
```
##Python
```{python}
boolean = data.rad > 2
data[boolean]
```
##R
```{r}
data %>%
  filter(data$zn > 25 & data$nox < 0.5)
```
##Python
```{python}
fil_1 = data.zn > 25
fil_2 = data.nox < 0.5
data[fil_1 & fil_2]
```

##R
```{r}
data %>%
  filter(zn < 50) %>%
  select(rad)
```
##Python
```{python}
data.rad[data.zn <50]
```

##R
```{r}
data$tant = data$zn + data$tax
head(data)
```
##Python
```{python}
data["tant"] = data.zn + data.tax
data.head()
```
```{python}
data.sort_values('tax',ascending = False)
```
##R
```{r}
library(corrplot)
nums <- sapply(data,is.numeric)
M <- cor(data[,nums])
corrplot.mixed(M)
```
##Python
```{python}
f,ax = plt.subplots(figsize = (18,18))
sns.heatmap(data.corr(),annot=True, linewidths=.5, fmt='.1f',ax=ax)
```
##R
```{r}
ggplot(data = data, aes(x=ptratio,group = 1))+
  geom_line(aes(y = black), linetype = "dashed", col = "red")+
  geom_line(aes(y = tax), linetype = "solid", color = "green")+
  labs(x = "X-axis", y = "Y-axis", title = "Line Plot")
```

##Python
```{python}
data.black.plot(kind = 'line', color = 'g', label = 'Blackness',linewidth=1, alpha=1, grid=True, linestyle=':')
data.tax.plot(color ='r', label='Tax amount', linewidth=1,alpha=1,grid=True, linestyle ='-.')
plt.legend(loc='upper right')
```

##R
```{r}
ggplot(data = data, aes(x = ptratio, y = tax, group = 1))+
  geom_point(col = "red") +
  labs(title = "Black tax scatter plot")
```
##Python
```{python}
data.plot(kind = 'scatter', x = 'ptratio', y = 'tax', alpha = 0.5, color = 'red')
plt.xlabel('Blackness')
plt.ylabel('Tax')
plt.title('Black tax scatter plot')
```

##R
```{r}
hist(data$tax, breaks = 100, main = 'Histogram of tax', col = 'blue')
```
```{python}
data.tax.plot(kind = 'hist', bins = 100, figsize = (5,5))
plt.show()
```

##R
```{r}
boxplot(data[,c(7:14)], las = 2)
```
##Python
```{python}
data.boxplot(column='tax', by = 'rad')
plt.show()
```
##R
```{r}
#Plotting all data
a <- ggplot(data = data, aes(x = ptratio)) +
  geom_line(aes(y = tax), col = "red")+
  geom_line(aes(y = black), col = "green")+
  geom_line(aes(y = tant), col = "blue")+
  labs(x = "X-axis",
       y = "Y-axis",
       title = "Line Plot")

b <- ggplot(data = data, aes(x = ptratio)) +
  geom_line(aes(y = tant), col = "red")

c <- ggplot(data = data, aes(x = ptratio))+
  geom_line(aes(y = black) , col = "green")

d <- ggplot(data = data, aes(x = ptratio))+
  geom_line(aes(y = tax), col = "blue")

library(gridExtra)

grid.arrange(a,b,c,d,nrow = 4, ncol = 1)
              
  
```
##Python
```{python}
data1 = data.loc[:,["tant","tax","black"]]
data1.plot()
data1.plot(subplots = True)
plt.show()
```

