---
title: "Support Vector Machine"
author: "Levi Orero"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Machine learning is a science of getting computers to act by feeding them data and letting them learn a few tricks on their own, without being explicitly programmed to do so.

_Supervised learning_
_Unsupervised learning_
_Reinforcement learning_

*Problem Statement*
To study a heart disease data set and to model a classifier for predicting whether a patient is suffering from any heart disease or not

#Load packages
```{r , message=FALSE}
library(caret)
library(readr)
```
#Import data
```{r}
heart <- read_csv("heart.csv")
```

```{r}
str(heart)
head(heart)
heart <- data.frame(heart)
```

#Split data
```{r}
intrain <- createDataPartition(y = heart$target, p = 0.7, list = FALSE)
training <- heart[intrain,]
testing <- heart[-intrain,]

#check sets
dim(training)
dim(testing)
                               
```
#Clean data
```{r}
anyNA(heart)
summary(heart)
```
```{r}
training[["target"]] = factor((training[["target"]]))
```

#train model
```{r}
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
#method defines resampling method
#number is the number of resampling iterations
#repeats contains sets to compute for our repeated cross validation

svm_Linear <- train(target ~.,data = training, method = "svmLinear",
                    trControl = trctrl, 
                    preProcess = c("center","scale"),
                    tunelength = 10)
                       
```
#check output
```{r}
svm_Linear
```

```{r}
#test data
test_pred <- predict(svm_Linear, newdata = testing)
test_pred
```

```{r}
#use confusion matrix to test accuracy of the model
confusionMatrix(table(test_pred, testing$target))
```
Output shows that our model accuracy is 80%
```{r}
grid <- expand.grid(C = c(0,0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1,
                          1.25, 1.5, 1.75, 2, 5))
svm_Linear_Grid <- train(target~., data = training, method = "svmLinear",
                        trControl = trctrl,
                        preProcess = c("center", "scale"),
                        tuneGrid = grid,
                        tuneLength = 10)
svm_Linear_Grid
plot(svm_Linear_Grid)
```
The above plot is showing that our clasifier is giving best accuracy on C = 0.1. Let's make predictions using this model for our test set.
```{r}
test_pred_grid <- predict(svm_Linear_Grid, newdata = testing)
test_pred_grid
```

Check accuracy using confusion matrix
```{r}
confusionMatrix(table(test_pred_grid, testing$target))
```

